---
title: "Grade Analysis Lab"
author: "Brian de Korodi"
format: html
editor: visual
keep-md: true
echo: FALSE
warning: false
---

## Question 1

```{r}
here::i_am("r-101-grade-analysis.Rproj")
here::i_am("grades.csv")
grades <- read.csv("grades.csv")
```

## Question 2

```{r}
exam_summary <-
  grades |>
  summarise("Mean grade" = mean(Exam, na.rm = TRUE),
                      "Median grade" = median(Exam, na.rm = TRUE),
                      "Lowest grade" = min(Exam, na.rm = TRUE),
                      "Highest grade" = max(Exam, na.rm = TRUE))
  
```

```{r}
knitr::kable(exam_summary, format = "markdown")
```

## Question 3

```{r}
missing_grades <- sum(is.na(grades$Exam))
```

```{r}
cat("There were", missing_grades,"students who did not take the final exam.")
```

## Question 4

```{r}
grades %>%
  ggplot(aes(x = Exam, na.rm = TRUE)) +
  geom_histogram(binwidth = 0.5, fill = "red") +
  labs(x = "Exam grades",
       y = "Frequency") +
  ggtitle("Distribution of exam grades")
```

## Question 5

```{r}
students_by_group <- 
  grades |>
  group_by(Group) |>
  summarise("Number_of_students" = n())
```

```{r}
knitr::kable(students_by_group)
```

## Question 6

```{r}
  students_by_group |>
    mutate('Group number' = gsub("grp_", "", students_by_group$Group)) |>
    ggplot(aes(x = `Group number`, y = `Number_of_students`)) +
    geom_bar(stat = "identity", fill = "red") +
    labs(title = "Number of students in each group",
         x = "Group number",
         y = "Number of students") 
```

## Question 7

One way to plot conditional distributions might be to plot the density functions for each group. Since there is a relatively large number of groups, it doesn't make sense to plot all the density functions on the same graph, so instead we can use a facet wrap. This visualization provides a solid quick overview of the shape of the distribution for each group:

```{r}
grades |>
    mutate('Group' = gsub("grp_", "", grades$Group)) |>
  ggplot(aes(x = `Exam`, color = "red")) + 
  geom_density(alpha = 0.5) +
  labs(title = "Distribution of grades by group") +
  facet_wrap(~Group)
```

A better solution might be to use a box plot. It allows us to plot the distributions for each group side-by-side, which enables easier comparison:

```{r}
grades |>
    mutate('Group number' = gsub("grp_", "", grades$Group)) |>
  ggplot(aes(x = `Group number`, y = `Exam`)) +
  geom_boxplot() +
  labs(title = "Distribution of grades by group",
       x = "Group number",
       y = "Grades")
```

The choice of which representation to use might depend on if one wants to compare between groups, or if one wants to simply be able to look up a specific group and get an idea for the general shape of its distribution.
